---
title: "GENFI_DataExploration"
author: "Shanny Foo"
date: "28/06/2022"
output: word_document
---

#### HOUSECLEANING IN R ####
# To comment/uncomment out whole chunk: SHIFT+CTRL+C
# To Restart R: SHIFT+CTRL+F10
```{r Clearing R Workspace, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# Clear plots
if(!is.null(dev.list())) dev.off()  # In GUI: clear all plots in the plots panel.

# Clear console
cat("\014")  # In GUI: Ctrl+L or to click on the clear console within the Edit menu.

# Clean workspace
rm(list=ls())  # In GUI: clear objects from the workspace in the environment panel.
```



--- STEP 1: SETUP

#### LIBRARIES ####
```{r warning=TRUE, include=FALSE}
#library(openxlsx)
library(readxl)

library(tidyverse)
library(ggpubr)
library(ggplot2)
```


#### SOURCE DATAFRAME: VARIABLE RE-NAMING AND ASSIGNMENT ####
```{r All Data, message=FALSE, warning=TRUE, include=FALSE, paged.print=FALSE}
## GENFI Data: Longitudinal by Rows: https://readxl.tidyverse.org/reference/read_excel.html
GENFI <- read_excel("E:\\Desktop\\FTD_Project\\GENFI\\GENFI_Data\\GENFI_Data_Analysis\\Dataframes\\GENFI_1_2_Longitudinal_03_19mar2022.xlsx", sheet = "Longitudinal_Rows")


## Mutate variables of interest
GENFI <- GENFI %>%
  mutate(Ant_Insula_Avg = (Ant_Insula_LEFT+Ant_Insula_RIGHT/2))


## Rename variables
GENFI <- GENFI %>%
  rename(#"Time" = Visit,
         "Subject" = Blinded_Code,
         "Days" = Date_of_assessment,
         "Age" = Age_at_visit,
         "ClinicalGroup" = Group_MC,
         "ClinicalStatus" = Group_MC_Coded,
         "GeneticMutation" = Genetic_Group_NMC,
         "GeneticGroup" = Genetic_Group_NMC_Coded,
         "MiniSEA" = MiniSEA_total,
         "RSMS" = RSMS_total,
         "Ekman" = Ekman_all,
         "IRI" = mIRI_total,
         "FauxPas" = Faux_pas_totalALL,
         "aINS" = Ant_Insula_Avg,
         "Site" = Blinded_Site_Coded,
         "ScannerMRI" = Scanner_Coded) %>%
    mutate_if(is.numeric,
            round,
            digits = 2)  # Round to 2 digits
```



--- STEP 2: DATA CLEANING AND PREPROCESSING

#### DF: VARIABLES OF INTEREST ####
```{r Labelling variables, echo=FALSE, message=FALSE, warning=TRUE, paged.print=FALSE}
## Timepoint 1
GENFI_Time_1 <- GENFI %>%
  #group_by(Subject) %>%
  select(Cohort, Days, Timepoint, Visit, Subject, Age, Gender, Education, ClinicalGroup, ClinicalStatus, GeneticMutation, GeneticGroup, MiniSEA, aINS, Site, ScannerMRI, Total_Brain, TIV_mm3) %>%
  filter(ClinicalGroup %in% c("NMC", "PMC", "SMC")) %>%
  filter(GeneticMutation %in% c("NMC", "C9", "GRN", "MAPT")) %>%
  filter(Timepoint == "1") 
```


#### IMPUTE TIV ####
```{r}
## Head size is unlikely to change within subject over time, so this value is imputed using median values across timepoints, regardless of NA

# Impute TIV_med
GENFI_Time_1 <- GENFI_Time_1 %>%
  #group_by(Subject) %>%
  mutate(TIV_med = median(TIV_mm3, na.rm = TRUE)) # Reduces number of missing subjects
  #ungroup()

# Remove TIV_mm3 and other unnecessary variables
GENFI_Time_1 <- GENFI_Time_1 %>% 
  select(-c(TIV_mm3, Cohort))
```



--- STEP 3: INSPECT AND EXPLORE THE DATA

#### INSPECTION OF COL FREQUENCY DISTRIBUTION ####
```{r Plot col freq dist, echo=FALSE, message=FALSE, warning=TRUE, paged.print=FALSE}
### GET A SENSE OF COL DISTRIBUTION

## Look at the summary statistics of the variables
summary(GENFI_Time_1)


### VISUAL INSPECTION OF COL FREQUENCY DISTRIBUTION

## Single plot of data distribution facted by variable
long <- reshape2::melt(GENFI_Time_1)

ggplot(long, aes(value)) +
      facet_wrap(~variable, scales = 'free_x') +
      geom_histogram(binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3)))
```


#### BOX PLOT: CATEGORICAL VARIABLE RANGES ####
```{r Plot categorial vars, echo=FALSE, message=FALSE, warning=TRUE, paged.print=FALSE}
## Check the range and distribution of the DV by categorical variables

par(mar=c(1,1,1,1))  # set plot margins

# ClinicalStatus and aINS
boxplot(aINS ~ ClinicalStatus, data = GENFI_Time_1)

# ClinicalStatus and MiniSEA
boxplot(MiniSEA ~ ClinicalStatus, data = GENFI_Time_1)

# ClinicalStatus and RSMS
boxplot(RSMS ~ ClinicalStatus, data = GENFI_Time_1)


# ClinicalStatus and GeneticMutation
boxplot(GeneticMutation ~ ClinicalStatus, data = GENFI_Time_1)


# ClinicalStatus and Age
boxplot(Age ~ ClinicalStatus, data = GENFI_Time_1)

# ClinicalStatus and Gender
boxplot(Gender ~ ClinicalStatus, data = GENFI_Time_1)

# ClinicalStatus and Education
boxplot(Education ~ ClinicalStatus, data = GENFI_Time_1)

# ClinicalStatus and TIV_mm3
boxplot(TIV_mm3 ~ ClinicalStatus, data = GENFI_Time_1)

# ClinicalStatus and Total_Brain
boxplot(Total_Brain ~ ClinicalStatus, data = GENFI_Time_1)

# ClinicalStatus and ScannerMRI
boxplot(ScannerMRI ~ ClinicalStatus, data = GENFI_Time_1)

# ClinicalStatus and Site
boxplot(Site ~ ClinicalStatus, data = GENFI_Time_1)

# ClinicalStatus and aINS (annualised change in GM volume)
boxplot(aINS ~ ClinicalStatus, data = GENFI_Time_1, cex=0.3)
```


#### SPAGHETTI PLOT: OVERALL DISTRIBUTION ####
```{r Plot overall dist, echo=FALSE, message=FALSE, warning=TRUE, paged.print=FALSE}
## Plot overall distribution of aINS over MiniSEA

# Label for plot title
lab.MiniSEA = c("11","12","13","14","15") # Label for MiniSEA to be added to ggplot

# Plot parameters
ggplot(data=GENFI_Time_1,
             aes(x=MiniSEA,y=aINS,group=Subject)) +
               geom_line () + geom_point () + ggtitle("aINS Spaghetti Plot") +
               theme_bw() +
               scale_x_continuous (breaks = c(11,12,13,14,15), labels=lab.MiniSEA, name = "MiniSEA") +
               scale_y_continuous (breaks = seq(-180,420,50), name= "aINS") +
               stat_summary(fun=mean,geom="point",size=2.5,shape=18,colour="orange") +
               stat_summary(fun=mean,geom="line",lwd=0.4,colour="blue") +
               stat_smooth(method="lm",se=FALSE,lwd=0.2,formula=y~x,colour="red") +
               stat_smooth(method="lm",se=FALSE,lwd=0.2,formula=y~poly(x,2),colour="green")
```


#### SPAGHETTI PLOT: CONDITIONING ON MiniSEA-INVARIANT COVARIATES ####
```{r Plot by group, echo=FALSE, message=FALSE, warning=TRUE, paged.print=FALSE}
## Spaghetti plot with lowess smoother by ClinicalStatus to get a sense of the mean function by ClinicalStatus

# Set panels (row#, col#), starting position (1,1,1,1 from margin), adjusted r, and font subtitle
par(mfrow=c(3,1),oma=c(1,1,1,1),adj=0.5,font.sub=3) # note: for error that panels were out of bounds, decreased start position to c(0,0,0,0)

# Plot parameters for each group
tmp1 = cbind(GENFI_Time_1$MiniSEA[GENFI_Time_1$ClinicalStatus=="0"],GENFI_Time_1$aINS[GENFI_Time_1$ClinicalStatus=="0"])

tmp2 = cbind(GENFI_Time_1$MiniSEA[GENFI_Time_1$ClinicalStatus=="1"],GENFI_Time_1$aINS[GENFI_Time_1$ClinicalStatus=="1"])

tmp3 = cbind(GENFI_Time_1$MiniSEA[GENFI_Time_1$ClinicalStatus=="2"],GENFI_Time_1$aINS[GENFI_Time_1$ClinicalStatus=="2"])

# Plot panels for each group, side by side
scatter.smooth(tmp1[,1],tmp1[,2],pch=42,cex=0.75,xlab="MiniSEA",
               ylab="aINS")
title("ClinicalGroup = NMC")

scatter.smooth(tmp2[,1],tmp2[,2],pch=42,cex=0.75,xlab="MiniSEA",
               ylab="aINS")
title("ClinicalGroup = PMC")

scatter.smooth(tmp3[,1],tmp3[,2],pch=42,cex=0.75,xlab="MiniSEA",
               ylab="aINS")
title("ClinicalGroup = SMC")
```


#### SCATTER PLOT: VARIABLE RELATIONSHIPS ####
```{r Plot x y relationship, echo=FALSE, message=FALSE, warning=TRUE, paged.print=FALSE}
## Plot the relationship between MiniSEA and aINS for ClinicalStatus groups, without taking into consideration the multilevel structure of the data
ggplot(data  = GENFI_Time_1,
       aes(x = MiniSEA,
           y = aINS))+
  geom_point(size = 1.2,
             alpha = .8,
             position = "jitter")+# to add some random noise for plotting purposes
  facet_wrap(~ClinicalStatus, nrow=4, ncol=1) +
  theme_minimal()+
  labs(title = "aINS vs. MiniSEA")


## Now, add a regression line to the plot
ggplot(data  = GENFI_Time_1,
       aes(x = MiniSEA,
           y = aINS))+
  geom_point(size     = 1.2,
             alpha    = .8,
             position = "jitter")+ #to add some random noise for plotting purposes
  geom_smooth(method = lm,
              se     = FALSE,
              col    = "black",
              size   = .5,
              alpha  = .8)+ # to add regression line
  facet_wrap(~ClinicalStatus, nrow=4, ncol=1) +
  theme_minimal()+
  labs(title    = "aINS vs. MiniSEA",
       subtitle = "add regression line")


## Now, draw different regression lines for the different ClinicalStatus groups in the same plot, faceted by GeneticMutation
ggplot(data      = GENFI_Time_1,
       aes(x     = MiniSEA,
           y     = aINS,
           col   = ClinicalStatus,
           group = ClinicalStatus))+ #to add the colours for different ClinicalStatus groups
  geom_point(size     = 1.2,
             alpha    = .8,
             position = "jitter")+ #to add some random noise for plotting purposes
  facet_wrap(~GeneticMutation, nrow=4, ncol=1) +
  theme_minimal()+
  theme(legend.position = "top")+
  #theme(legend.position = "none")+
  scale_fill_steps2()+
  #scale_color_gradientn(colours = rainbow(100))+
  geom_smooth(method = lm,
              se     = FALSE,
              size   = .5,
              alpha  = .8)+ # to add regression line
  labs(title    = "aINS vs. MiniSEA",
       subtitle = "add colours for different ClinicalStatus groups and regression lines")
```