---
title: "Calculate amount of missing scans by subject"
output: word_document
editor_options: 
  chunk_output_type: console
---

# Load data
# mutate variable to count number of 1's and then subtract the 1's from each download missing and GENFI missing rows. 
# treat the beh missing manually since there are only 2

#### LIBRARIES ####
```{r}
library(openxlsx)
library(readxl)
library(writexl)

library(tidyverse)

```


#### SOURCE DATAFRAME ####
```{r All Data, message=FALSE, warning=TRUE, include=FALSE, paged.print=FALSE}
## GENFI Data: Longitudinal by Rows: https://readxl.tidyverse.org/reference/read_excel.html
GENFI <- read_excel("E:\\Desktop\\FTD_Project\\GENFI\\GENFI_Data\\GENFI_Data_Analysis\\Dataframes\\GENFI_1_2_Longitudinal_04_07jun2022.xlsx", sheet = "Longitudinal_Rows")
```


#### COUNT MISSING SCANS ####
```{r}
## Code variables to count missing scans: Code missing as 1, otherwise 0
GENFI$Scans_Missing_Coded <- ifelse(GENFI$Scans_Missing == "MISSING", 1, 0)
#GENFI$Scans_Missing_Coded <- as.numeric(GENFI$Scans_Missing_Coded)
GENFI$BIDS_anat_Missing <- ifelse(GENFI$BIDS_anat == "MISSING", 1, 0)
GENFI$BIDS_func_Missing <- ifelse(GENFI$BIDS_func == "MISSING", 1, 0)


## Create dataframe of just missing scans
GENFI_Missing <- GENFI %>%
  select(Group_MC, Genetic_Group_NMC, Cohort, Blinded_Site, Scanner, Date_of_assessment, Session, Visit, Blinded_Code, BIDS_anat, BIDS_func, Download_Missing, GENFI_Missing, Scans_Missing, Beh_Missing, BIDS_anat_Missing, BIDS_func_Missing, Scans_Missing_Coded, Row_Count) %>%
  filter(Scans_Missing == "MISSING" | Beh_Missing == "MISSING" | BIDS_anat == "MISSING" | BIDS_func == "MISSING") %>%
  group_by(Blinded_Code) %>%
  mutate(Scans_Expected = sum(Row_Count),
         Files_Missing = sum(Scans_Missing_Coded),
         Scans_Count = Scans_Expected-Files_Missing,
         anat_expected = sum(Row_Count),
         func_expected = sum(Row_Count),
         anat_missing = sum(BIDS_anat_Missing),
         func_missing = sum(BIDS_func_Missing),
         anat_count = anat_expected-anat_missing,
         func_count = func_expected-func_missing)


## Code variables to count how many subjects should have at least 2 timepoints
GENFI_Missing$anat_long <- ifelse(GENFI_Missing$anat_expected>1, 1, 0)
GENFI_Missing$func_long <- ifelse(GENFI_Missing$func_expected>1, 1, 0)

## Create dataframe to find out who are these subs with affected longitudinal
GENFI_Long_Missing <- GENFI_Missing %>%
  select(Group_MC, Genetic_Group_NMC, Blinded_Code, anat_expected, func_expected, anat_missing, func_missing, anat_long, func_long) %>%
  filter(func_expected>1 | anat_expected>1) 
  #filter(func_expected-func_missing < 2 | anat_expected-anat_missing < 2)

```


#### WRITE TO EXCEL ####
```{r}
# Source: https://datatofish.com/export-dataframe-to-excel-in-r/

# library("writexl")
# 
# df <- data.frame(Name = c("Jon", "Bill", "Maria", "Ben", "Tina"),
#                  Age = c(23, 41, 32, 58, 26)
#                  )

write_xlsx(GENFI_Missing, "E:\\Desktop\\FTD_Project\\GENFI\\GENFI_Data\\GENFI_Missing_Count.xlsx")
```


#### SUMMARY: MISSING FOLDERS ####
```{r}
## Subject scan file folders

# Total rows: 69
GENFI_Missing %>%
  ungroup() %>%
  filter(Group_MC == "SMC") %>%
  filter(Scans_Count<Scans_Expected) %>%
  summarise(count=n()) 

# Total subjects: 59
GENFI_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "SMC") %>%
  filter(Scans_Count<Scans_Expected) %>%
  summarise(count=n())

###

# Total rows: 81
GENFI_Missing %>%
  ungroup() %>%
  filter(Group_MC == "PMC") %>%
  filter(Scans_Count<Scans_Expected) %>%
  summarise(count=n())

# Total subjects: 60
GENFI_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "PMC") %>%
  filter(Scans_Count<Scans_Expected) %>%
  summarise(count=n())

###

# Total rows: 105
GENFI_Missing %>%
  ungroup() %>%
  filter(Group_MC == "NMC") %>%
  filter(Scans_Count<Scans_Expected) %>%
  summarise(count=n())

# Total subjects: 62
GENFI_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "NMC") %>%
  filter(Scans_Count<Scans_Expected) %>%
  summarise(count=n())

```


#### SUMMARY: MISSING ANAT AND FUNC ####
```{r}
## anat

# Total rows: 71
GENFI_Missing %>%
  ungroup() %>%
  filter(Group_MC == "SMC") %>%
  filter(anat_count<anat_expected) %>%
  summarise(count=n())

# Total subjects: 61
GENFI_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "SMC") %>%
  filter(anat_count<anat_expected) %>%
  summarise(count=n())

###

# Total rows: 83
GENFI_Missing %>%
  ungroup() %>%
  filter(Group_MC == "PMC") %>%
  filter(anat_count<anat_expected) %>%
  summarise(count=n())

# Total subjects: 61
GENFI_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "PMC") %>%
  filter(anat_count<anat_expected) %>%
  summarise(count=n())

###

# Total rows: 111
GENFI_Missing %>%
  ungroup() %>%
  filter(Group_MC == "NMC") %>%
  filter(anat_count<anat_expected) %>%
  summarise(count=n())

# Total subjects: 68
GENFI_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "NMC") %>%
  filter(anat_count<anat_expected) %>%
  summarise(count=n())

####################

## func

# Total rows: 125
GENFI_Missing %>%
  ungroup() %>%
  filter(Group_MC == "SMC") %>%
  filter(func_count<func_expected) %>%
  summarise(count=n())

# Total subjects: 110
GENFI_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "SMC") %>%
  filter(func_count<func_expected) %>%
  summarise(count=n())

###

# Total rows: 242
GENFI_Missing %>%
  ungroup() %>%
  filter(Group_MC == "PMC") %>%
  filter(func_count<func_expected) %>%
  summarise(count=n())

# Total subjects: 170
GENFI_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "PMC") %>%
  filter(func_count<func_expected) %>%
  summarise(count=n())

###

# Total rows: 229
GENFI_Missing %>%
  ungroup() %>%
  filter(Group_MC == "NMC") %>%
  filter(func_count<func_expected) %>%
  summarise(count=n())

# Total subjects: 150
GENFI_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "NMC") %>%
  filter(func_count<func_expected) %>%
  summarise(count=n())

```


#### SUMMARY: SUBS MISSING LONG THIS CODE NEEDS FIXING ####
```{r}
## anat_long_missing

# Total rows: 
GENFI_Missing %>%
  ungroup() %>%
  filter(Group_MC == "SMC") %>%
  filter(anat_long_missing == 1) %>%
  summarise(count=n())

# Total subjects: 
GENFI_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "SMC") %>%
  filter(anat_long_missing ==1) %>%
  summarise(count=n())

###

# Total rows: 
GENFI_Missing %>%
  ungroup() %>%
  filter(Group_MC == "PMC") %>%
  filter(anat_long_missing == 1) %>%
  summarise(count=n())

# Total subjects: 
GENFI_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "PMC") %>%
  filter(anat_long_missing ==1) %>%
  summarise(count=n())

###

# Total rows: 
GENFI_Missing %>%
  ungroup() %>%
  filter(Group_MC == "NMC") %>%
  filter(anat_long_missing == 1) %>%
  summarise(count=n())

# Total subjects: 
GENFI_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "NMC") %>%
  filter(anat_long_missing ==1) %>%
  summarise(count=n())

##############################

## func_long_missing

# Total rows: 
GENFI_Missing %>%
  ungroup() %>%
  filter(Group_MC == "SMC") %>%
  filter(func_long_missing == 1) %>%
  summarise(count=n())

# Total subjects: 
GENFI_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "SMC") %>%
  filter(func_long_missing ==1) %>%
  summarise(count=n())

###

# Total rows: 
GENFI_Missing %>%
  ungroup() %>%
  filter(Group_MC == "PMC") %>%
  filter(func_long_missing == 1) %>%
  summarise(count=n())

# Total subjects: 
GENFI_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "PMC") %>%
  filter(func_long_missing ==1) %>%
  summarise(count=n())

###

# Total rows: 
GENFI_Missing %>%
  ungroup() %>%
  filter(Group_MC == "NMC") %>%
  filter(func_long_missing == 1) %>%
  summarise(count=n())

# Total subjects: 
GENFI_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "NMC") %>%
  filter(func_long_missing ==1) %>%
  summarise(count=n())

```


#### SUMMARY: CHECK FOR UNIQUE SUBS ####
```{r}
# # List of unique subs to check if missing data affects longitduinal, sorted alphabetically
# GENFI_Missing_Unique <- sort(unique(GENFI_Missing$Blinded_Code))
# 
# # Convert list to dataframe
# GENFI_Missing_Unique <- as.data.frame(GENFI_Missing_Unique)


# Total rows: 36
GENFI_Long_Missing %>%
  ungroup() %>%
  filter(Group_MC == "SMC") %>%
  summarise(count=n())

# Total subjects: 21
GENFI_Long_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "SMC") %>%
  summarise(count=n())

### 

# Total rows: 135
GENFI_Long_Missing %>%
  ungroup() %>%
  filter(Group_MC == "PMC") %>%
  summarise(count=n())

# Total subjects: 22
GENFI_Long_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "PMC") %>%
  summarise(count=n())

###

# Total rows: 130
GENFI_Long_Missing %>%
  ungroup() %>%
  filter(Group_MC == "NMC") %>%
  summarise(count=n())

# Total subjects: 51
GENFI_Long_Missing %>%
  group_by(Blinded_Code) %>%
  filter(Group_MC == "NMC") %>%
  summarise(count=n())

```






